BHRC Preliminary Analyses

### 2021-10-26 BHRC_Prelim_Analyses.Rmd
Added visualizations

---
title: "BHRC Preliminary Analyses"
output: html_notebook
---

@@

combined.data2<-merge(proposal.data.w0,sle_factors,by="ident")
all.data<-merge(combined.data1,combined.data2,by="SUBJECTID")
```

### Remove test set:
```{r}
set.seed(2021)
test=sample(seq(1015),200,replace=FALSE)
test.data<-all.data[test,]
all.data<-all.data[-test,]
```

@@

all.data.complete.set <- all.data[(all.data$ALERT>(ALERT.mean-ALERT.3sd) & all.data$ALERT<(ALERT.mean+ALERT.3sd)) & (all.data$ORIENT>(ORIENT.mean-ORIENT.3sd) & all.data$ORIENT<(ORIENT.mean+ORIENT.3sd)) & (all.data$CONFLICT>(CONFLICT.mean-CONFLICT.3sd) & all.data$CONFLICT<(CONFLICT.mean+CONFLICT.3sd)),]
```

### Run linear models using Threat and Deprivation scores to predict ANT performance metrics
```{r}
summary(lm(ALERT~age+gender+threat+depriv,data=all.data.ALERT))
summary(lm(ORIENT~age+gender+threat+depriv,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+threat+depriv,data=all.data.CONFLICT))

summary(lm(ALERT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv,data=all.data.ALERT)) # sig gender x threat
summary(lm(ORIENT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv,data=all.data.CONFLICT)) # close age x threat

summary(lm(ALERT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv+age*gender*threat+age*gender*depriv,data=all.data.ALERT))
summary(lm(ORIENT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv+age*gender*threat+age*gender*depriv,data=all.data.ORIENT)) 
summary(lm(CONFLICT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv+age*gender*threat+age*gender*depriv,data=all.data.CONFLICT))

summary(lm(ALERT~age+gender+depriv+depriv*age+depriv*gender,data=all.data.ALERT)) 
summary(lm(ORIENT~age+gender+depriv+depriv*age+depriv*gender,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+depriv+depriv*age+depriv*gender,data=all.data.CONFLICT))

summary(lm(ALERT~age+gender+threat+threat*age+threat*gender,data=all.data.ALERT)) # Close on gender x threat
summary(lm(ORIENT~age+gender+threat+threat*age+threat*gender,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+threat+threat*age+threat*gender,data=all.data.CONFLICT))
```

@@

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(depriv,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$age)+geom_smooth(method=lm)+ facet_wrap(~ factor(gender,labels=c("Gender 1","Gender 2")))

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(depriv,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$gender)+geom_smooth(method=lm)+ facet_grid(factor(all.data.ALERT[!is.na(all.data.ALERT$gender),]$gender,labels=c("Gender 1","Gender 2"))~factor(round(all.data.ALERT[!is.na(all.data.ALERT$gender),]$age)))

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(threat,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$age)+geom_smooth(method=lm)+ facet_wrap(~ factor(gender,labels=c("Gender 1","Gender 2")))

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(threat,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$gender)+geom_smooth(method=lm)+ facet_grid(factor(gender,labels=c("Gender 1","Gender 2"))~factor(round(age)))

@@

model1<-lm(ALERT~age+gender,data=all.data.ALERT)
model2<-lm(ALERT~age+gender+threat,data=all.data.ALERT)
model3<-lm(ALERT~age+gender+threat+threat*age+threat*gender,data=all.data.ALERT)
model4<-lm(ALERT~age+gender+threat+threat*age+threat*gender+age*gender*threat,data=all.data.ALERT)
model5<-lm(ALERT~age+gender+threat+threat*gender,data=all.data.ALERT)
anova(model1,model2,test = "Chisq")
anova(model2,model3,test = "Chisq")
anova(model3,model4,test = "Chisq")
anova(model2,model5,test = "Chisq") # close 0.057
```

### What if we split by symptoms?

@@

summary(lm(threat~age+ALERT+ORIENT+CONFLICT,data=all.data.complete.set[all.data.complete.set$gender==1,]))
summary(lm(depriv~age+ALERT+ORIENT+CONFLICT,data=all.data.complete.set[all.data.complete.set$gender==1,]))

summary(lm(threat~age+ALERT+ORIENT+CONFLICT,data=all.data.complete.set[all.data.complete.set$gender==2,]))
summary(lm(depriv~age+ALERT+ORIENT+CONFLICT,data=all.data.complete.set[all.data.complete.set$gender==2,]))

>>>>

---
title: "BHRC Preliminary Analyses"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=250)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lsr)
```

@@

combined.data2<-merge(proposal.data.w0,sle_factors,by="ident")
all.data<-merge(combined.data1,combined.data2,by="SUBJECTID")
all.data$gender<-as.factor(all.data$gender)
```

### Remove test set:
```{r}
set.seed(2021)
test=sample(seq(1015),200,replace=FALSE)
test.data<-all.data[test,] # Test set (n=200)
all.data<-all.data[-test,] # Training set (n=815)
```

@@

# This is a complete set with subjects who have all 3 measures within the bounds of 3SD
all.data.complete.set <- all.data[(all.data$ALERT>(ALERT.mean-ALERT.3sd) & all.data$ALERT<(ALERT.mean+ALERT.3sd)) & (all.data$ORIENT>(ORIENT.mean-ORIENT.3sd) & all.data$ORIENT<(ORIENT.mean+ORIENT.3sd)) & (all.data$CONFLICT>(CONFLICT.mean-CONFLICT.3sd) & all.data$CONFLICT<(CONFLICT.mean+CONFLICT.3sd)),]
```

### Threat and deprivation scores are correlated, so we'll test them in separate models
```{r}
cor.test(all.data.complete.set$threat,all.data.complete.set$depriv)
```

### Run linear models using Threat and Deprivation scores to predict ANT performance metrics
# Note: Gender 1 = Male, Gender 2 = Female
```{r}
# Without interactions
summary(lm(ALERT~age+gender+threat,data=all.data.ALERT))
summary(lm(ORIENT~age+gender+threat,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+threat,data=all.data.CONFLICT))

summary(lm(ALERT~age+gender+depriv,data=all.data.ALERT))
summary(lm(ORIENT~age+gender+depriv,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+depriv,data=all.data.CONFLICT))

# With interactions
summary(lm(ALERT~age+gender+depriv+depriv*age+depriv*gender,data=all.data.ALERT)) 
summary(lm(ORIENT~age+gender+depriv+depriv*age+depriv*gender,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+depriv+depriv*age+depriv*gender,data=all.data.CONFLICT))

summary(lm(ALERT~age+gender+threat+threat*age+threat*gender,data=all.data.ALERT)) # Close on gender x threat
summary(lm(ORIENT~age+gender+threat+threat*age+threat*gender,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+threat+threat*age+threat*gender,data=all.data.CONFLICT))

summary(lm(ALERT~age+threat,data=all.data.ALERT[all.data.ALERT$gender==1,]))
summary(lm(ALERT~age+threat,data=all.data.ALERT[all.data.ALERT$gender==2,])) # significant just in girls

# Including threat and deprivation together in the same model (just to see?)
summary(lm(ALERT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv,data=all.data.ALERT)) # sig gender x threat
summary(lm(ORIENT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv,data=all.data.CONFLICT))

@@

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(depriv,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$age)+geom_smooth(method=lm)+ facet_wrap(~ factor(gender,labels=c("Male","Female")))

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(depriv,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$gender)+geom_smooth(method=lm)+ facet_grid(factor(all.data.ALERT[!is.na(all.data.ALERT$gender),]$gender,labels=c("Male","Female"))~factor(round(all.data.ALERT[!is.na(all.data.ALERT$gender),]$age)))

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(threat,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$age)+geom_smooth(method=lm)+ facet_wrap(~ factor(gender,labels=c("Male","Female")))

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(threat,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$gender)+geom_smooth(method=lm)+ facet_grid(factor(gender,labels=c("Male","Female"))~factor(round(age)))

@@

model1<-lm(ALERT~age+gender,data=all.data.ALERT)
model2<-lm(ALERT~age+gender+threat,data=all.data.ALERT)
model3<-lm(ALERT~age+gender+threat+threat*age+threat*gender,data=all.data.ALERT)
model4<-lm(ALERT~age+gender+threat+threat*age+threat*gender+age*gender*threat,data=all.data.ALERT)
model5<-lm(ALERT~age+gender+threat+threat*gender,data=all.data.ALERT)
anova(model1,model2,test = "Chisq")
anova(model2,model3,test = "Chisq")
anova(model3,model4,test = "Chisq")
anova(model2,model5,test = "Chisq") # close 0.057

### Next steps: What if we split by symptoms?

@@
